% \documentclass[12pt]{scrreprt}
\documentclass[12pt]{report} 

% language may be romanian or english (default is english)
% type may be bachelor or master (default is bachelor)
\usepackage[language=english, type=bachelor]{style}
\usepackage{graphicx}   % for images
\usepackage{caption}    % for captionof

%\geometry{a4paper,top=2.5cm,left=3cm,right=2.5cm,bottom=2.5cm}
%in style
%controlling the appearance of your headers and footers
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\chead{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.2pt}

\hypersetup{hidelinks}

\begin{document}

\specialization{INFORMATICĂ ENGLEZĂ}	
\title{Doclane}					   
\author{Robert Béres}											
\supervisor{Lect. PETRESCU Manuela}				
				
\maketitle


\newpage
\thispagestyle{empty}
\mbox{}
\newpage
\pagenumbering{roman} 

\cleardoublepage
ABSTRACT
\vspace{0.5cm}	
\hrule
\vspace{0.5cm}	
%\cleardoublepage

Document-based administrative workflows in public institutions remain largely dependent on physical presence, paper submissions, and manual processing. Citizens are required to visit offices, wait in queues, and resubmit documents when something is missing or incorrect, often without any visibility into the status of their request between visits. While digital tools exist in this space, they tend to cover only part of the process, handle a single service type, or require significant custom development before an institution can use them.

This thesis presents Doclane, a web-based platform designed to digitise the full document request lifecycle between citizens and public institutions. The system allows citizens to browse configurable request templates, upload required documents in structured slots, and track the status of their submission in real time. Institution employees can claim incoming requests, review documents individually, and communicate feedback directly through the platform when corrections are needed. The workflow continues iteratively until all documents are approved, reducing repeated physical visits to only those required by law, such as the final collection of an issued document.

Doclane is built on a Go backend organised into clearly separated handler, service, repository, and model layers, and a Next.js frontend using server-side rendering to minimise load times on data-heavy pages. The system is deployed on AWS using a containerised, Kubernetes-based architecture, with infrastructure managed entirely through Terraform. Authentication is handled by AWS Cognito, file storage by S3 with presigned URLs, and document processing is augmented by three AWS managed services: Textract for OCR-based text extraction, Bedrock for LLM-assisted document validation using the Amazon Nova Lite model, and Polly for text-to-speech synthesis. No custom model training was performed; these services are consumed through their managed APIs.

The system is designed to be institutionally configurable rather than institution-specific. Request templates, required document slots, and department structures are all defined through the admin interface, meaning the same deployment can serve different institutional contexts without changes to the underlying codebase. This distinguishes Doclane from bespoke administrative software, which requires separate development for each institution, and from large national platforms, which are out of reach for smaller public institutions with limited technical resources.

The thesis also reviews related work in electronic document management and e-government systems, drawing on academic literature as well as real deployments such as Estonia's X-Road and the UK's GOV.UK platform, and situates Doclane relative to the EU's eIDAS regulatory framework for electronic identification and trust services.

\tableofcontents


\newpage
\pagenumbering{arabic}

\input{chapters/chapter1}
\input{chapters/chapter2}
\input{chapters/chapter3}
\input{chapters/chapter4}
\input{chapters/chapter5}
\input{chapters/chapter6}
\input{chapters/chapter7}

\bibliography{references}

\end{document}
